Visualizing probability distributions across bivariate cyclic temporal granularities

### 2020-03-12 main.Rmd
elements of grammar explaination cut short, levels cut short

The grammar of graphics introduced a framework to construct statistical graphics by relating data space to the graphic space [@Wilkinson1999-nk]. The layered grammar of graphics proposed by [@Wickham2009pk], which is an alternate and modified parametrization of the grammar suggests that graphics are made up of distinct layers of grammatical elements. There are seven grammatical elements in total, out of which data, aesthetics and geometries are essential elements. Data are observations made on a number of variables, typically thought of as a sample from a population distribution. Aesthetic mapping describes how variables are mapped to elements of the plot, such as horizontal axis or color. Geometries determine the physical representation of the data in the plot. The remaining grammatical elements of a graphic namely facets, statistics, coordinates and themes are optional and contains detail of the plot. Facets dictates how to split up our plot and is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data and can see whether they are the same or different.

@@

Drawing from the grammar of graphics, if <$C_1$, $C_2$, $v$> serves as the basis of visualizing the distribution of the measured variable, the following layers can be specified:

- Data: <$C_1$, $C_2$, $v$>
- Aesthetic mapping: $C_1$ mapped to $x$ position and $v$ to $y$ position
- Statistical transformation: Any descriptive or smoothing statistics that summarizes distribution of $v$
- Geometric objects: Any geometry displaying distribution, for example, boxplot, letter value, violin, ridge or highest density region plots
- Facet:  $C_2$

@@

The following principles are useful while choosing distribution plots given two temporal granularities:

  - If levels of both granularity plotted are low/medium, then any distribution plots might be chosen depending on which feature of the distribution needs focus.

 - If level of the granularity plotted across x-axis is more than medium, ridge plots should be avoided to escape overlap of categories.

  - If level of the granularity plotted across x-axis is more than or equal to high, quantile plots are preferred.

  - If levels of any granularity plotted are more than medium, quantile based methods of visualizing distribution is preferred as they use less space by design than most density based methods.

>>>>

The grammar of graphics introduced a framework to construct statistical graphics by relating data space to the graphic space [@Wilkinson1999-nk]. The layered grammar of graphics proposed by [@Wickham2009pk], which is an alternate and modified parametrization of the grammar suggests that graphics are made up of distinct layers of grammatical elements. Drawing from the grammar of graphics, if <$C_1$, $C_2$, $v$> serves as the basis of visualizing the distribution of the measured variable, the following layers can be specified:

<!-- CONTENT CUT -->
<!-- There are seven grammatical elements in total, out of which data, aesthetics and geometries are essential elements. Data are observations made on a number of variables, typically thought of as a sample from a population distribution. Aesthetic mapping describes how variables are mapped to elements of the plot, such as horizontal axis or color. Geometries determine the physical representation of the data in the plot. The remaining grammatical elements of a graphic namely facets, statistics, coordinates and themes are optional and contains detail of the plot. Facets dictates how to split up our plot and is a powerful tool for exploratory data analysis as we can rapidly compare patterns in different parts of the data and can see whether they are the same or different. -->

@@

- Data: <$C_1$, $C_2$, $v$>
- Aesthetic mapping (mapping of variables to elements of the plot): $C_1$ mapped to $x$ position and $v$ to $y$ position
- Statistical transformation (summarize data): Any descriptive or smoothing statistics that summarizes distribution of $v$
- Geometric objects (physical representation of the data): Any geometry displaying distribution, for example, boxplot, letter value, violin, ridge or highest density region plots
- Facet (split plots):  $C_2$

@@

<!-- CONTENT CUT -->
<!-- The following principles are useful while choosing distribution plots given two temporal granularities: -->

<!--   - If levels of both granularity plotted are low/medium, then any distribution plots might be chosen depending on which feature of the distribution needs focus. -->

<!--  - If level of the granularity plotted across x-axis is more than medium, ridge plots should be avoided to escape overlap of categories. -->

<!--   - If level of the granularity plotted across x-axis is more than or equal to high, quantile plots are preferred. -->

<!--   - If levels of any granularity plotted are more than medium, quantile based methods of visualizing distribution is preferred as they use less space by design than most density based methods. -->
